\documentclass[american]{article}
\usepackage{lmodern}
\usepackage[latin9]{inputenc}
\usepackage[a4paper]{geometry}
\geometry{verbose,tmargin=1cm,bmargin=4cm,lmargin=3cm,rmargin=3cm}
\usepackage{color}
\usepackage{amstext}
\usepackage[authoryear]{natbib}

\makeatletter
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}

\makeatother

\usepackage{babel}
\usepackage{listings}
\lstset{basicstyle={\footnotesize\bfseries\ttfamily},
breaklines=true,
commentstyle={\itshape\color{OliveGreen}},
keywordstyle={\bfseries\color{blue}},
showstringspaces=true,
stringstyle={\bfseries\color{RoyalPurple}}}
\begin{document}
\title{Advanced radiation and remote sensing }
\author{Manfred Brath, Oliver Lemke, Stefan Bühler}
\maketitle

\subsection*{Exercise No. 4 -- Atmospheric Brightness Temperature Spectra}
The function ``run\_arts" within the Jupyter Notebook \texttt{rtcalc.ipynb} simulates 
a brightness temperature and a atmospheric zenith opacity spectrum in the 
microwave spectral range for a midlatitude-summer atmosphere over a smooth and wet
land surface for a desired height and looking direction.
\begin{enumerate}
\item  
Run the function ``run\_arts" with the given values for height and direction.
Ignore the brightness temperature for now and consider the zenith opacity spectrum 
to answer the following questions:
\begin{itemize}
\item The spectrum includes four spectral lines. To which species do these
lines belong? Play around with the absorption species selection in
the Jupyter Notebook.
\item We speak of window regions where the zenith opacity is below 1. Where
are they?
\end{itemize}
\item Brightness temperature is a unit for intensity. It is the temperature
of a blackbody that emits the same amount of intensity.
Mathematically, the transformation between intensity in SI units and
intensity in brightness temperature is done with the Planck formula.
ARTS is capable to perform simulation in units of brightness temperature.
Uncomment the code part for the second task. Investigate the brightness
temperature spectra for different hypothetical sensors:
\begin{enumerate}
\item A ground-based sensor looking in the zenith direction.
\item A sensor on an airplane ($z=10\,\text{km}$) looking in the zenith
direction.
\end{enumerate}
Consider both opacity and brightness temperatures to answer the following
questions: 
\begin{itemize}
\item In plot (a), why do the lines near 60 GHz and near 180 GHz appear
flat on top? 
\item In plot (b), why is the line at 180 GHz smaller than before? 
\item Describe the difference between plots (a) and (b). What happens to
the lines, what happens to the background? Can you explain what you
see? 
\end{itemize}
\item Make the same calculation as in task 2 for a satellite sensor ($z=800\,\text{km}$)
looking nadir (straight down).

Answer following questions: 
\begin{itemize}
\item Explain the brightness temperature simulated in the window regions. 
\item Why does the line at 22 GHz look different from the others? 
\item Investigate the the $\text{O}_{2}$ line at $120\,\text{GHz}$. Perform
an ARTS simulation focused around that frequency. Why does the shape
close to the center of the $\text{O}_{2}$ line at $120\,\text{GHz}$
looks so differently compared to the $183\,\text{GHz}$. 
\end{itemize}
\end{enumerate}

\end{document}
